---
title: "Data wrangling"
author: "Bibechana Paudel"
date: "2025-03-19"
output: 
  md_document:
      variant: gfm
---

### Data wrangling is manipulating the data that are usually large and messy. 
### Required packages
```{r}
#### Load the packages
library(tidyverse)
```

### Load the data
```{r}
microbiome.fungi<-read.csv("Data wrangling/Bull_richness.csv", na.strings="na")
str(microbiome.fungi)

```
### 'select()'
We can use select to only select the columns that we require. So here we are selecting the columns we prefer. ":" will select all the columns from Compartment upto Fungicide.  
```{r}
microbiome.fungi2<-select(microbiome.fungi, SampleID, Crop, Compartment:Fungicide, richness)
```
### filter()
This is a convenient function that allows us to subset data easily. 
```{r}
head(filter(microbiome.fungi2, Treatment=="Conv."))

# A more complex using &
head(filter(microbiome.fungi2, Treatment=="Conv." & Fungicide=="C"))

# Another more complex example using or |
head(filter(microbiome.fungi2, Sample=="A" | Sample=="B")) # sample A or B

```
### mutate()
Allows us to quickly create new columns. 
```{r}
#microbiome.fungi2$logRich<-log(microbiome.fungi2$richness)  ### In basic R we did this
# Create a new column called logRich
head(mutate(microbiome.fungi2, logRich=log(richness)))   

# Create a new column which combines Crop and Treatment
head(mutate(microbiome.fungi2, Crop_Treatment=paste(Crop, Treatment)))

```
### the pipe '%>%'
We can combine the output from one function to input for another function. It helps us to visualise what we are doing to our data in each step. Basically the pipe allows us to input the previous input data from the right side of the pipe to the function on the left. 
```{r}
microbiome.fungi %>%  
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% #selecting columns
  filter(Treatment=="Conv.") %>% #subsetting to only include the conventional treatment
  mutate(logRich=log(richness)) %>% # creating a new columnto the log richness
    head() # displaying the first six rows
```
### summarise()
We use the summarise to find means, standard deviations or standard errors. 
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich)) # calculating overall mean log richness within the conventionally managed treatment. Here the Mean.rich is the column name
```

We can also connect multiple summaries. 
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  filter(Treatment == "Conv.") %>% # subsetting to only include the conventional treatment
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(),  # calculates the number of rows
            sd.dev = sd(logRich)) %>%     # calculate the sd of log of Richness
  mutate(std.err = sd.dev/sqrt(n))  # add a new column with std.error
```

## group_by()
We can use this function to summarise statistics by group.
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n))
```

### Connecting to plotting
These groups of functions and packages work together very nicely. Direct input into ggplot is possible. In this case, we would not have to enter the data statement; we just need to start inputting your aesthetics.
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  mutate(logRich = log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.rich = mean(logRich), # calculating the mean richness, stdeviation, and standard error
            n = n(), 
            sd.dev = sd(logRich)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  ggplot(aes(x = Fungicide, y = Mean.rich)) + # adding in a ggplot
  geom_bar(stat="identity") +   # create a bar graph
  geom_errorbar( aes(x=Fungicide, ymin=Mean.rich-std.err, ymax=Mean.rich+std.err), width=0.4) +
  theme_minimal() +
  xlab("") +
  ylab("Log Richness") +
  facet_wrap(~Treatment)
```

### Joining
The join group of functions are super useful. Have you ever tried to match up data by rows based on some criteria? It's almost impossible to mess up over 10-15 rows of data.

More info can be found here: https://dplyr.tidyverse.org/reference/mutate-joins.htmlLinks to an external site.

But we can easily do this with the joining functions. They include:

- left_join() - Keep all rows of X and add matching rows from Y. Any rows in Y that don’t match X are excluded.

- right_join() - reverse of left_join() 
-inner_join() - only keep rows that are common to both X AND Y, remove everything else. 
- full_join() - Keep any columns that are in either X or Y

To try these out, let's split our data into separate sets, including taxonomy and metadata.
Here we are also sampling 100 random rows of our dataset using the 'sample_n()' function
```{r}
# selecting just the richness and sample ID
richness <- microbiome.fungi %>%
  select(SampleID, richness)

# selecting columns that don't include the richness
metadata <- microbiome.fungi %>% 
  select(SampleID, Fungicide, Crop, Compartment, GrowthStage, Treatment, Rep, Sample)

head(metadata)
head(richness)

head(left_join(metadata, richness, by = "SampleID")) # adding the richness data to the metadata based on on the common column of sampleID

```

### Pivoting
Pivoting is also useful for converting from wide to long format and back again. We can do this with pivot_longer() and pivot_wider()

More info can be found here: https://tidyr.tidyverse.org/reference/pivot_wider.htmlLinks to an external site.

I will demonstrate pivot_wider() assuming that most data is already formatted as long data. Lets explore turning our data set into a wide format, since its in long format.

```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) # calculates the mean per Treatment and Fungicide 
```

Wide format - sets the values within the fungicide column into column names
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) # pivot to wide format. We want the name of fungicide to be put into the new columns and values are from mean. 
```

Easily can take the difference bewteen the fungicide and control now.
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  filter(Class == "Sordariomycetes") %>%
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) # calculate the difference between the means. 
```

Now we can easily calculate the difference between fungicide and control and plot it.
```{r}
microbiome.fungi %>%
  select(SampleID, Crop, Compartment:Fungicide, richness) %>% # selecting columns  
  group_by(Treatment, Fungicide) %>% # grouping by treatment and fungicide to later calculate summary stats by group
  summarise(Mean = mean(richness)) %>% # calculates the mean per Treatment and Fungicide 
  pivot_wider(names_from = Fungicide, values_from = Mean) %>% # pivot to wide format
  mutate(diff.fungicide = C - F) %>%  # calculate the difference between the means. 
  ggplot(aes(x = Treatment, y = diff.fungicide)) + # Plot it 
  geom_col() +
  theme_minimal() +
  xlab("") +
  ylab("Difference in average species richness")
```

