---
title: "Homework Iteration and Function"
author: "Bibechana Paudel"
date: "2025-03-25"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
  md_document:
    variant: gfm
---

### Function to convert Fahrenheit to celsius
```{r}
#(5*(degree_f-32)/9)
(5*(32-32)/9)
### Creating a function 
F_to_C<-function(f_temp){     ### the ()has the variable     
  ### {}has the code
 celsius<-(5*(f_temp-32)/9) ### saving the output as the Celsius
 return(celsius)    ###returns output
}

F_to_C(32)   ### after creating the function we can run the code as number R function with the desired variable. 
F_to_C(-40)
```

### Function to convert celsius to  Fahrenheit
```{r}
### Write a function to convert celsius to fahrenheiht using the conversion of F = C x (9/5)+32
C_to_F<-function(celsius_temp){
  fahrenheit<-(celsius_temp*(9/5)+32)
  return(fahrenheit)
}
C_to_F(-40)
C_to_F(32)
```


### Iteration function in base R
Iterations are useful for reproducibility, helps to reduce copy and paste errors.
```{r}
rep("A",3)   ### rep function for repeat: repeat A 3 times
rep(c("A","B"),10) ###repeat A and B ten times
rep(c(1,2,5,3),4, each=5) ### repeat 1,2,5,and 3 4 times and 5 times each.
```

### seq() function 
Helps towrite sequence of numbers
```{r}
1:7
seq(from=1, to=7) #sequence of number from 1 to 7.
seq(from=0, to=10, by=2)  ##sequence of numbers from 0 to 10 by 2s.
```

### seq_along() 
To generate a sequence of numbers based on non-integer values.

```{r}
# use the built in LETTERS vector for an example.
LETTERS
seq_along(LETTERS)   
```


### The for loop
```{r}
for(i in 1:10){
 print(i*2 )    ### for each value of i(1 to 10), it gives the output using the 2 times of i until it reaches the 10. 
}

```


### Combining functions and iterations
```{r}
for(i in -30:100){   ###for every value of i from -30 to 100, it converts the Fahrenheit to celsius and print the result. 
  result<-F_to_C(i)
  print(result)
}
###But we cannot save those output as they are in the console and we need to copy and paste that.
```
Step 1. Set a R object to NULL

Step 2. Set your for loop

Step 3. Save the result of your for loop into a dataframe each iteration

Step 4. append one row of the dataframe to the null object each iteration of the loop. 

```{r}
celsius.df<-NULL
for(i in -30:100){
  result<-data.frame(F_to_C(i), i)    ##it creates a dataframe called as result with two columns and one column will have the fahrenheit and the other with the celsius converted value
  print(result)
  
  celsius.df<-rbind.data.frame(celsius.df,result)   ### rbind will bind the results to new iterations.  It basically binds the previous iterations to new one. 
}
celsius.df
```

### Challenge example
Load the data and required library
```{r}
EC50.data<-read.csv("Iteration_function/EC50_all.csv",na.strings="na")
library(tidyverse)
# install.packages("drc")
library(drc)
str(EC50.data)
```


```{r}
isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == "ILSO_5-41c"] ~ 
        EC50.data$conc[EC50.data$is == "ILSO_5-41c"], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    
```

### The loop
```{r}
###unique function find the unique value of the character 
nm<-unique(EC50.data$is)     ###seq along the character vector nm.
for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
  print(nm[[i]])
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    EC50
}
```


```{r}
EC50.ll4<-NULL
nm<-unique(EC50.data$is)     ###seq along the character vector nm.
for (i in seq_along(nm)) {
  isolate1 <- drm(100 * EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~ 
        EC50.data$conc[EC50.data$is == nm[[i]]], 
                       fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                  names = c("Slope", "Lower", "Upper", "EC50")), 
                       na.action = na.omit)
  print(nm[[i]])
    # outputs the summary of the paramters including the estimate, standard
    # error, t-value, and p-value outputs it into a data frame called
    # summary.mef.fit for 'summary of fit'
    summary.fit <- data.frame(summary(isolate1)[[3]])
    # outputs the summary of just the EC50 data including the estimate, standard
    # error, upper and lower bounds of the 95% confidence intervals around the
    # EC50
    EC50 <- ED(isolate1, respLev = c(50), type = "relative", 
        interval = "delta")[[1]]
    isolate.ec_1<-data.frame(nm[[i]], EC50)
    EC50.ll4<-rbind.data.frame(EC50.ll4,isolate.ec_1)
    EC50
}
```


### Loop using tidyverse
```{r}
EC50.data %>%
  group_by(is) %>%     ####group by the isolate column
  nest() %>%      #### nest function helps to create a data frame within the dataframe. 
  mutate(ll.4.mod = map(data, ~drm(.$relgrowth ~ .$conc,     ####create a new column called LL.4 , map function tells that we want to  (it is for loop contained within the function), period(.) symbol inherits all the columns from the from the data. The ‘.’ tells R that we are considering all the data. Then a $ followed by the column headers in the data as the input into our drm function.
                              fct = LL.4(fixed = c(NA, NA, NA, NA), 
                                         names = c("Slope", "Lower", "Upper", "EC50"))))) %>%
  ### we want to mat the function over the model to scrub the EC50 value. 
  mutate(ec50 = map(ll.4.mod, ~ED(., 
                              respLev = c(50),    
                              type = "relative",
                              interval = "delta")[[1]])) %>%
  unnest(ec50) ###unnest will  ‘extract’ the value from the nesting. 
```

